---
title: "Data Import"
output: github_document
---

This file is for doing data import

```{r}
library(tidyverse)
library(readxl)
library(haven)
```
R wants to know where the files are (eg. where is the excel file?)

Two ways to do this:

- *Absolute path*: the full path from the "north pole"
- *Relative path*: path from where you are right now. Easy if you are doing the rdirectory because it can be reproducible. ALWAYS use relative paths. 

Import our first data set
```{r}
litters_df =
  read_csv(file = "data/FAS_litters.csv",
           na = c("", "NA", "."))
#if there is something that looks like numbers but is assigned character, something is wrong
#check the CSV and you will see that they used . and NA to represent NA values
#to change this, you can rename the NA categories
#there are many ways to data clean using read_csv

litters_df = janitor :: clean_names(litters_df)

#this makes the names in snake case rather than with a space 
```

Import your second data set
```{r}
pups_df =
  read_csv(
    file= "data/FAS_pups.csv", 
    skip = 3,
    na = c("", "NA", ".")) # this part allows you to skip the first three rows 

pups_df = janitor:: clean_names(pups_df) 
#to prevent yourself to load too many libraries, you can use this format to use one function from a library without loading the entire thing. The package needs to be installed though. 

```
Head or tail 
```{r}
#look at the first rows of the data set
head(litters_df, 20) #number tells you how many rows to show

#look at the last rows of the data set 
tail(litters_df,5)
```

Skimming 
```{r}
skimr::skim(pups_df)
```

Viewing
```{r}
#view(pups_df)
#don't do this just do it in the console because otherwise every time you knit it will pop up an 
#additional window to view the df
```

reading csv!

Make sure you are using read_csv because read.csv is super old 

## Oh excel...


Jenny Bryan made 'readxl' to solve our problems
```{r}
mlb11 =
  read_excel("data/mlb11.xlsx")

```

Look at the data
```{r}
mlb11
```

Load some LotR data

Import FOTR words
```{r}
fotr_df=
  read_excel(
    "data/LotR_Words.xlsx", 
    range = "B3:D6" #top left corner to bottom right
  )

#look at ?read_excel to see how to specify the range 
```

What about the two towers?
```{r}
tt_df=
  read_excel(
    "data/LotR_Words.xlsx", 
    range = "F3:H6" #top left corner to bottom right
  )

tt_df
```

If you have excel file open while you are working in RStudio, it will pop up in your git window. It will disappear if you close excel. Never commit these. 

##Import SAS

Read in the PULSE data set

```{r}
#similar structure to the others
pulse_df=
  read_sas("data/public_pulse_data.sas7bdat")

pulse_df = janitor:: clean_names(pulse_df)

#don't make variables capital, makes your variables follow a clean naming
#convention
```

